---
title: "EDA_consulting"
author: "Maggie Sha"
date: "4/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(stringi)
```


1. Demographic data of respondents (INCLUDE only those who selected Public School)
 a. World Language Teacher or Administrator
 b. Language(s) Taught
 c. Level(s) Taught (elem, middle, high)

2. Response to question: Has your school or district established proficiency targets for each of the levels of world language study? 
 a. This is asked in the Outcome section for elem, middle, and high
 b. Please report on the response by level

3. Internal Consistency of Teacher Practices:  start from 1610
 a. Check the internal consistency of scores in the Teacher Practices section
  i. Cronbach Alpha coefficient
  ii. See spreadsheet for sub-scales
  iii. Please note certain items need to be reversed when checking for internal consistency (these are noted as “negative” in the spreadsheet)


```{r}
#read file

data_fpath <- file.path(
  "data", 
  "World Language Programming Survey in MA - Data to MSSP 6-10-2020.xlsx"
)

first_two_rows <- read_excel(data_fpath, n_max = 1)

qnums <- first_two_rows %>%
  names() %>%
  janitor::make_clean_names()

qnames <- first_two_rows %>%
  flatten_chr()

data <- read_excel(data_fpath, skip = 2, col_names = qnames) %>%
  arrange(`End Date`) %>%
  mutate(response_num = row_number()) %>%
  rename(district =
           starts_with("Please select your school district"),
         languages_taught =
           matches("Which languages do you currently teach.+ choice$"))

data_clean_names <- data %>%
  janitor::clean_names()

#keep columns that I will use
data <- data_clean_names[,c(9:11, 13)]
```


```{r}
#add clean_title(administrator/teacher) column

# some titles are hard to determine: "specialist", "liaison" "supervisor"
# use answers in language_taught as the second filter
administrator_title <- c("head", "chair", "leader", "chairperson", "director", "supervisor", "administrator")
data$clean_title <- "Teacher"


# create data frame with teach/administrator title and the corresponding count
for(i in 1:nrow(data)){
  detect <- sum(stri_detect_fixed(tolower(data$part_i_demographics_your_title)[i], administrator_title))
  if(detect >= 1){
    data$clean_title[i] <- "Administrator"
  } else if(grepl("chair/administrator/supervisor", data$languages_taught[i])){
    data$clean_title[i] <- "Administrator"
  }
}

Title.df <- data.frame(table(data$clean_title))
colnames(Title.df) <- c("Title","Count")
```

#1.a
```{r}
#visualize title
ggplot(Title.df, aes(x = Title, y = Count, fill = Title)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(Count/sum(Count),digits = 2)), vjust = 2, color= "black", size=3)
```
```{r}
# clean columns 'languages_taught' and 'what_level_s_do_you_currently_teach_please_check_all_that_apply'
# remove the answer "I am a World Language Program chair/administrator/supervisor"

stat <- "I am a World Language Program chair/administrator/supervisor"

data2 <- data # for language question
data3 <- data # for level question

for(i in 1:nrow(data2)){
  detect2 <- sum(stri_detect_fixed(data2$languages_taught[i], stat))
  if(detect2 >= 1) {
    x <- unlist(strsplit(data2$languages_taught[i], " "))
    data2$languages_taught[i] <- ifelse((c(x[1],x[2]) == c("I", "am")), 
                                      str_remove(data2$languages_taught[i], "I am a World Language Program chair/administrator/supervisor"), 
                                      str_remove(data2$languages_taught[i], ",I am a World Language Program chair/administrator/supervisor"))
    }
}
  
for(i in 1:nrow(data3)){
  detect3 <- sum(stri_detect_fixed(data3$what_level_s_do_you_currently_teach_please_check_all_that_apply[i], stat))
  if(detect3 >= 1) {
    y <- unlist(strsplit(data3$what_level_s_do_you_currently_teach_please_check_all_that_apply[i], " "))
    
    if (c(y[1],y[2]) == c("I", "am")){
      data3$what_level_s_do_you_currently_teach_please_check_all_that_apply[i] <- 
        str_remove(data3$what_level_s_do_you_currently_teach_please_check_all_that_apply[i],
                   "I am a World Language Program chair/administrator/supervisor")
    } else{
      data3$what_level_s_do_you_currently_teach_please_check_all_that_apply[i] <- 
        str_remove(data3$what_level_s_do_you_currently_teach_please_check_all_that_apply[i],
                   ",I am a World Language Program chair/administrator/supervisor")
    }
  }
}

```


#1.b
```{r}
# create data frame with each kind of language and the corresponding count
# a person will be counted as 1 in all language levels that this person taught

language.df <- data.frame(language = character(), count = numeric())
data2$languages_taught[data2$languages_taught ==""] <- NA
data2 <- na.omit(data2)

for(i in 1:nrow(data2)){
  split <- unlist(strsplit(data2$languages_taught[i], ","))
  ind <- length(split) 
   for (j in 1:ind){
     lan <- split[j]
     if(!split[j] %in% language.df$language){
       # print(i)
       # print(split[j])
       language.df <- rbind.data.frame(language.df, cbind.data.frame(language = lan, count = 1))
     } else{
       ind2 <- which(language.df$language == lan)
       language.df[ind2,]$count <- language.df[ind2,]$count + 1
     }
   }
}
    
# visualize language_taught distribution
ggplot(language.df, aes(x = language, y = count, fill = language)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.9)) +
  geom_text(aes(label = count), vjust = 2, color= "black", size=3) + 
  theme(axis.text.x = element_text(angle = 60)) + 
  labs(title = "Languages taught by each individual - aggregated")
```


#1.c
```{r}
# create data frame with teach levels (ele,mid,high) and the corresponding counts
# a person will be counted as 1 in all levels that this person taught

level.df <- data.frame(level = character(), count = numeric())
data3$what_level_s_do_you_currently_teach_please_check_all_that_apply[data3$what_level_s_do_you_currently_teach_please_check_all_that_apply == ""] <- NA
data3 <- na.omit(data3)


for(i in 1:nrow(data3)){
  split <- unlist(strsplit(data3$what_level_s_do_you_currently_teach_please_check_all_that_apply[i], ","))
  ind <- length(split) 
   for (j in 1:ind){
     lev <- split[j]
     if(!lev %in% level.df$level){
       # print(i)
       # print(split[j])
       level.df <- rbind.data.frame(level.df, cbind.data.frame(level = lev, count = 1))
     } else{
       ind2 <- which(level.df$level == lev)
       level.df[ind2,]$count <- level.df[ind2,]$count + 1
     }
   }
}

# only keep level = ele, mid, and high
ele_mid_high <- level.df[c(1,2,4),]

# visualize teach level distrubution
ggplot(ele_mid_high, aes(x = level, y = count, fill = level)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.9)) +
  geom_text(aes(label = count), vjust = 2, color= "black", size=3) + 
  theme(axis.text.x = element_text(angle = 20)) +
  labs(title = "Level taught by each individual - aggregated")
```

#2
```{r}
# For each district:
# * If one or more administrators, include the administrator with greater progress
# * If more than one administrators have the same progress, take the first administrator
# * If no admin, take the first teacher to respond to the survey

one_per_district <- data_clean_names %>%
  select(response_num, district, progress, duration_in_seconds, finished,
         languages_taught,has_your_school_or_district_established_proficiency_targets_for_each_of_the_levels_of_world_language_study_249,
         has_your_school_or_district_established_proficiency_targets_for_each_of_the_levels_of_world_language_study_742,
         has_your_school_or_district_established_proficiency_targets_for_each_of_the_levels_of_world_language_study_1291) %>%
  mutate(is_admin = str_detect(languages_taught,
                               "I am a World Language Program chair")) %>%
  group_by(district) %>%
  filter(
    case_when(
      any(is_admin && finished)  ~ is_admin,
      !any(is_admin) ~ response_num == min(response_num)
    )
  ) %>%
  ungroup()

one_per_district <- one_per_district %>% group_by(district) %>%
  filter(progress == max(progress)) %>% filter(response_num == min(response_num))
```

#2.b
Response to question: Has your school or district established proficiency targets for each of the levels of world language study?

```{r}
# distribution of responses by elementary school, middle school, and high school level

ele_targets <- data.frame(table(one_per_district$has_your_school_or_district_established_proficiency_targets_for_each_of_the_levels_of_world_language_study_249))
ele_targets <- ele_targets %>% filter(Var1 != "Not sure")
colnames(ele_targets) <- c("Answer","count")

mid_targets <- data.frame(table(one_per_district$has_your_school_or_district_established_proficiency_targets_for_each_of_the_levels_of_world_language_study_742))
mid_targets <- mid_targets %>% filter(Var1 != "Not sure")
colnames(mid_targets) <- c("Answer","count")

high_targets <- data.frame(table(one_per_district$has_your_school_or_district_established_proficiency_targets_for_each_of_the_levels_of_world_language_study_1291))
high_targets <- high_targets %>% filter(Var1 != "Not sure")
colnames(high_targets) <- c("Answer","count")


ele <- ggplot(ele_targets, aes(x = Answer, y = count, fill = Answer)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.9)) +
  geom_text(aes(label = count), vjust = 2, color= "black", size=3) + 
  theme(axis.text.x=element_blank()) +
  labs(title = "Elementary Level")

mid <- ggplot(mid_targets, aes(x = Answer, y = count, fill = Answer)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.9)) +
  geom_text(aes(label = count), vjust = 2, color= "black", size=3) + 
  theme(axis.text.x=element_blank()) +
  labs(title = "Middle School Level")

high <- ggplot(high_targets, aes(x = Answer, y = count, fill = Answer)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.9)) +
  geom_text(aes(label = count), vjust = 2, color= "black", size=3) + 
  theme(axis.text.x=element_blank()) +
  labs(title = "High School Level")

ele
mid
high
```